---
title: rmmc
subtitle: An R package for calculating multimarket contact

# Summary for listings and search engines
summary: "Over the last few months I have worked on creating a package that can help with calculating various measures of multimarket contact. This is a short post to illustrate how it can be used."

# Link this post with a project
projects:
- content/project/rmmc

# Date published
date: "2021-11-29T00:00:00Z"

# Date updated
lastmod: "2021-11-29T00:00:00Z"

# Is this an unpublished draft?
draft: false

# Show this page in the Featured widget?
featured: false

# Featured image
# Place an image named `featured.jpg/png` in this page's folder and customize its options here.
# image:
#  caption: 'Image credit: [**Unsplash**](https://unsplash.com/photos/CpkOjOcXdUY)'
#  focal_point: ""
#  placement: 2
#  preview_only: false

authors:
- admin


tags:
- multimarket contact

categories: ""

output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Multimarket contact is the situation when firms meet in more than one market. To capture this situation in empirical research a number of different measures have been proposed. The R package `rmmc` can be used as a central repository of functions that makes it easy to access the various functions that have been used. 

This post provides a short illustration about how this package can be used. 

First, install start by installing the package. Currently is only available on github. 

```{r}
library(devtools)
install_github("grlju/rmmc")
```

Let's load some example data provide with the package

```{r}
library(rmmc)
data <- rmmc:::data
head(data)
```

The `data.table` (the package `data.table` package is a dependency) as information on three firms (firm1-firm3) across years (2000-2003) and four markets in which these firms have activities (m1-m4). This example dataset also illustrates the layout that the package expects. If there are multiple observations for each entity the data should be in long format. 

Calculating the multimarket contact between these firms can now easily be accomplished with the various functions. 

Let's use a simple count measure. There can be multiple level at which we can think that multimarket contact exists and is important. Let's focus on the dyadic level first. That is, the level of contact between two different entities (i.e. firms in our example). 

```{r}
count_dyad <- mmc_count(data, firm_col = "id", date_col = "year", level = "dyad", market_cols = c("m1", "m2", "m3", "m4"))
head(count_dyad)
```

This calculates a count of the dyadic contact between the firms in data. For instance, firm1 and firm2 have no contact in year 2000 but firm1 and firm3 have 2 contacts in the same year.

